Local PCA Shows How Population Structure Differs Along the Genome

### 2016-06-28 writeup/local_pca_paper.tex
update

\begin{figure}
    \begin{center}
       \includegraphics{Fig1_allchr_Together_MDS_plot_compact}
    \end{center}
    \caption{
        Variation in structure for windows across \textit{Drosophila melanogaster} chromosome arms.
        In all plots, each point represents one window along the genome.
         The first column shows the MDS visualization of relationships between windows,
         and the second and third columns show the position (midpoint) of each window along the genome
         against the first and second MDS coordinates, respectively. 
         Rows show results from different chromosome arms, from top to bottom these are 2L, 2R, 3L, 3R, and X. 
         Colors are consistent for plots in each row. 
         Vertical black lines show the breakpoints of known polymorphic inversions
         (from top to bottom: In(2L)t, In(2R)NS, In(3L)OK, In(3R)K).
         \label{fig:mds_chr2L}
    }
\end{figure}

@@

The variation in population structure turns out to be almost entirely explained by
several large inversions that are polymorphic in these samples, identified in \citet{corbett2012population}. 
% Frequencies of each inversion in the sample are given in Table \ref{tab:inversion_freqs}.
We recolored the PCA plots in Figure \ref{fig:pca_by_pop} by the orientation of the inversion for each sample,
obtaining Figure \ref{fig:pca_by_inversion}.
Taking chromosome arm 2L as an example,
the two regions of similar, extreme population structure
shown in green in the first row of Figure \ref{fig:mds_chr2L}
lie directly around the breakpoints of the iversion In(2L)t,
and the PCA plots in the first rows of figure \ref{fig:pca_by_inversion}
shows that population structure here is mostly determined by inversion orientation.
The regions shown in purple lie near the centromere,
and have population structure reflective of two axes of variation,
seen in figure \ref{fig:pca_by_pop},
which correspond roughly to latitude within Africa and to degree of cosmopolitan admixture respectively
(see \citet{pool2015natural} for more details).
\plr{check that reference}
The regions shown in orange mostly lie inside the inversion,
and show population structure that is a mixture between the other two,
as expected due to recombination within the (long) inversion \citep{kirkpatrick2015chromosome}.
\plr{check that reference too}
Similar results are found in other chromosome arms,
\plr{add chromosome X}
except that the situation of Chr3R is a little more complicated,
due to the coexistence of two polymorphic inversions (In(3R)K and In(3R)P) at intermediate frequency.
However, peaks in MDS coordinates along the chromosome show the breakpoints of both,
although we have not highlighted portions of the MDS plot that pick these out.
Two further inversions, In(3L)P and In(3R)Mo, are at very low frequency in these samples, 
and do not significantly influence population structure.

\begin{figure}
    \begin{center}
       \includegraphics{Fig3_pca_plots_color_by_inv_allchr}
    \end{center}
    \caption{
         As in Figure 2, except that samples are colored by orientation of the corresponding polymorphic inversions, 
         In(2L)t, In(2R)NS, In(3L)OK and In(3R)K respectively. 
         (Data from \citet{lack2015drosophila}.
        \label{fig:pca_by_inversion}
    }
\end{figure}

>>>>

\begin{figure}
    \begin{center}
       \includegraphics{Fig1_allchr_Together_MDS_plot_compact_with_ChrX_inv}
    \end{center}
    \caption{
        Variation in structure for windows across \textit{Drosophila melanogaster} chromosome arms.
        In all plots, each point represents one window along the genome.
         The first column shows the MDS visualization of relationships between windows,
         and the second and third columns show the position (midpoint) of each window along the genome
         against the first and second MDS coordinates, respectively. 
         Rows show results from different chromosome arms, from top to bottom these are 2L, 2R, 3L, 3R, and X. 
         Colors are consistent for plots in each row. 
         Vertical black lines show the breakpoints of known polymorphic inversions
         (from top to bottom: In(2L)t, In(2R)NS, In(3L)OK, In(3R)K).
         \label{fig:mds_chr2L}
    }
\end{figure}

@@

The variation in population structure turns out to be almost entirely explained by
several large inversions that are polymorphic in these samples, identified in \citet{corbett2012population}. 
% Frequencies of each inversion in the sample are given in Table \ref{tab:inversion_freqs}.
We recolored the PCA plots in Figure \ref{fig:pca_by_pop} by the orientation of the inversion for each sample,
obtaining Figure \ref{fig:pca_by_inversion}.
Taking chromosome arm 2L as an example,
the two regions of similar, extreme population structure
shown in green in the first row of Figure \ref{fig:mds_chr2L}
lie directly around the breakpoints of the inversion In(2L)t,
and the PCA plots in the first rows of figure \ref{fig:pca_by_inversion}
shows that population structure here is mostly determined by inversion orientation.
The regions shown in purple lie near the centromere,
and have population structure reflective of two axes of variation,
seen in figure \ref{fig:pca_by_pop},
which correspond roughly to latitude within Africa and to degree of cosmopolitan admixture respectively
(see \citet{pool2015natural} for more details).
\plr{check that reference}
The regions shown in orange mostly lie inside the inversion,
and show population structure that is a mixture between the other two,
as expected due to recombination within the (long) inversion \citep{kirkpatrick2015chromosome}.
\plr{check that reference too}
Similar results are found in other chromosome arms,
\plr{add chromosome X}
except that the situation of Chr3R is a little more complicated,
due to the coexistence of two polymorphic inversions (In(3R)K and In(3R)P) at intermediate frequency.
However, peaks in MDS coordinates along the chromosome show the breakpoints of both,
although we have not highlighted portions of the MDS plot that pick these out.
Two further inversions, In(3L)P and In(3R)Mo, are at very low frequency in these samples, 
and do not significantly influence population structure.

\begin{figure}
    \begin{center}
       \includegraphics{Fig3_pca_plots_color_by_inv_allchr_with_ChrX_In1A}
    \end{center}
    \caption{
         As in Figure 2, except that samples are colored by orientation of the corresponding polymorphic inversions, 
         In(2L)t, In(2R)NS, In(3L)OK and In(3R)K respectively. 
         (Data from \citet{lack2015drosophila}.
        \label{fig:pca_by_inversion}
    }
\end{figure}
